### 2021-01-05 thesis/thesis.tex
move paper into folder as thesis, so I can write up all that I want to write

% Papers should be formatted according to the two-column ACM proceedings
% format.  Each paper should have no more than 12 pages, excluding
% bibliography, in 10pt font. There is no limit on the page count for
% references. Each reference must list all authors of the paper (do not use et
% al). The citations should be in numeric style, e.g., [52]. Submissions should
% be in PDF format and printable on US Letter and A4 sized paper. These
% requirements are all the same as in the previous year.

\def\paperversiondraft{draft}
\def\paperversionblind{blind}
\def\paperversioncamera{camera}

% If no special paper-version is requested, compile in draft mode
\ifx\paperversion\paperversionblind
\else
  \ifx\paperversion\paperversioncamera
  \else
     \def\paperversion{draft}
  \fi
\fi

\def\grammarlyon{on}

\ifx\grammarly\grammarlyon
\def\review{}
\else
\def\review{review,}
\fi

\ifx\paperversion\paperversiondraft
  \documentclass[sigplan,\review anonymous]{acmart}
\fi

\ifx\paperversion\paperversionblind
  \documentclass[sigplan,\review anonymous]{acmart}
\fi

\ifx\paperversion\paperversioncamera
  \documentclass[sigplan]{acmart}\settopmatter{}
\fi

% Most PL conferences are edited by conference-publishing.com. Follow their
% advice to add the following packages.
%
% The first enables the use of UTF-8 as character encoding, which is the
% standard nowadays. The second ensures the use of font encodings that support
% accented characters etc. (Why should I use this?). The mictotype package
% enables certain features 'to­wards ty­po­graph­i­cal per­fec­tion
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}

\usepackage{xargs}
\usepackage{lipsum}
\usepackage[textsize=tiny]{todonotes}
\usepackage{xparse}
\usepackage{xifthen, xstring}
\usepackage{ulem}
\usepackage{xspace}
\usepackage{minted}
\usemintedstyle{bw}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue}
\usepackage{mathpartir}
\usepackage{dirtytalk}
\usepackage[ligature, inference]{semantic}

\usepackage{listings}
\usepackage{xcolor}

%New colors defined below
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

%Code listing style named "mystyle"
% https://www.overleaf.com/project/5fc73ba8b2793a82677493d5
% https://tex.stackexchange.com/questions/200208/how-can-i-reduce-the-line-spacing-in-a-listing
\lstdefinestyle{mystyle}{
  commentstyle=\color{gray},
  keywordstyle=\bfseries,
  numberstyle=\ttfamily\footnotesize\color{codegray},
  basicstyle=\linespread{0.8}\ttfamily,%,\footnotesize,
  breakatwhitespace=false,
  breaklines=true,
  captionpos=b,
  keepspaces=true,
  numbers=left,
  columns=flexible,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=2
}

%"mystyle" code listing set
\lstset{style=mystyle}

% Use to keep figure in same section
% \usepackage[section]{placeins}

\makeatletter
\font\uwavefont=lasyb10 scaled 652

\newcommand\colorwave[1][blue]{\bgroup\markoverwith{\lower3\p@\hbox{\uwavefont\textcolor{#1}{\char58}}}\ULon}
\makeatother

\ifx\paperversion\paperversiondraft
\newcommand\createtodoauthor[2]{%
\def\tmpdefault{emptystring}
\expandafter\newcommand\csname #1\endcsname[2][\tmpdefault]{\def\tmp{##1}\ifthenelse{\equal{\tmp}{\tmpdefault}}
   {\todo[linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1:} ##2}}
   {\ifthenelse{\equal{##2}{}}{\colorwave[#2]{##1}\xspace}{ \todo[linecolor=#2,backgroundcolor=#2,bordercolor=#2]{\textbf{#1:} ##2}\colorwave[#2]{##1}}}}}
\else
\newcommand\createtodoauthor[2]{%
\def\tmpdefault{emptystring}
\expandafter\newcommand\csname #1\endcsname[2][\tmpdefault]{\def\tmp{##1}\ifthenelse{\equal{\tmp}{\tmpdefault}}
   {}
   {\ifthenelse{\equal{##2}{}}{##1\xspace}{ ##1}}}}
\fi

% Broaden margins to make room for todo notes
\ifx\paperversion\paperversiondraft
  \paperwidth=\dimexpr \paperwidth + 6cm\relax
  \oddsidemargin=\dimexpr\oddsidemargin + 3cm\relax
  \evensidemargin=\dimexpr\evensidemargin + 3cm\relax
  \marginparwidth=\dimexpr \marginparwidth + 3cm\relax
  \setlength{\marginparwidth}{4.6cm}
\fi

% We use the following color scheme
% 
% This scheme is both print-friendly and colorblind safe for
% up to four colors (including the red tones makes it not
% colorblind safe any more)
%
% https://colorbrewer2.org/#type=qualitative&scheme=Paired&n=4

\definecolor{pairedOneLightBlue}{HTML}{a6cee3}
\definecolor{pairedTwoDarkBlue}{HTML}{1f78b4}
\definecolor{pairedThreeLightGreen}{HTML}{b2df8a}
\definecolor{pairedFourDarkGreen}{HTML}{33a02c}
\definecolor{pairedFiveLightRed}{HTML}{fb9a99}
\definecolor{pairedSixDarkRed}{HTML}{e31a1c}

\createtodoauthor{grosser}{pairedOneLightBlue}
% \createtodoauthor{authorTwo}{pairedTwoDarkBlue}
\createtodoauthor{sid}{pairedThreeLightGreen}
\createtodoauthor{authorFour}{pairedFourDarkGreen}
\createtodoauthor{authorFive}{pairedFiveLightRed}
\createtodoauthor{authorSix}{pairedSixDarkRed}

%% Note: Authors migrating a paper from traditional SIGPLAN
%% proceedings format to PACMPL format should change 'sigplan' to
%% 'acmsmall'.

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{marvosym}

% [Operational semantics](https://tex.stackexchange.com/questions/271147/how-to-align-the-premises-in-a-semantic)
\renewcommand{\inference}[3][]{%
  \[\begin{array}[b]{@{}lc}
      \\
      \begin{array}[b]{l} #2 \end{array}
      \smash{\raisebox{-0.6\normalbaselineskip}{\scriptsize}} \\
      % \cline{2-2}
      \cmidrule[0.4pt]{1-1}
      \begin{array}[t]{l} #3 \end{array}
      \\
      \\
  \end{array}\]
}

% https://tex.stackexchange.com/questions/198462/the-lightning-symbol-in-an-equation-environment
%\newcommand{\lightning}{\mbox{\Lightning}}
\newcommand{\val}{\texttt{value} }
\newcommand{\unk}{\texttt{unk} }
\newcommand{\absent}{\texttt{absent} }
\newcommand{\case}{\texttt{case} }
\newcommand{\CONS}{\texttt{cons} }
\newcommand{\lzap}{\texttt{lz.ap} }
\newcommand{\lzthunk}{\texttt{lz.thunk} }
\newcommand{\construct}{\texttt{lz.construct} }
\newcommand{\stdcall}{\texttt{std.call} }
\newcommand{\std}{\texttt{std} }
\newcommand{\nofib}{\texttt{nofib} }
\newcommand{\Int}{\texttt{Int} }

\newcommand{\cpp}[1]{\mintinline{cpp}{#1}}
\newcommand{\raw}[1]{\mintinline{text}{#1}}
\newcommand{\denotation}[1]{\left\llbracket {\raw{#1}} \right\rrbracket}

\newcommand{\mlir}{\texttt{MLIR} }
\newcommand{\lz}{\texttt{lz} }
\newcommand{\lzforce}{\texttt{lz.force}}
\newcommand{\lzconstruct}{\texttt{lz.construct}}
\newcommand{\GHC}{\texttt{GHC} }
\newcommand{\ghc}{\GHC}
\newcommand{\Core}{\core}
\newcommand{\core}{\texttt{Core} }

% syntactic elements of our language
\newcommand{\EXPR}{\texttt{EXPR}}
\newcommand{\PHI}{\texttt{PHI}}
\newcommand{\BB}{\texttt{BB}}
\newcommand{\BBID}{\texttt{BBID}}
\newcommand{\INST}{\texttt{INST}}
\newcommand{\TERM}{\texttt{TERM}}
\newcommand{\LABEL}{\texttt{LABEL}}
\newcommand{\FLAG}{\texttt{FLAG}}
\newcommand{\ASSIGN}{\texttt{ASSIGN}}
\newcommand{\VID}{\texttt{VID}}

% transition relations
\newcommand{\assignarrow}{\xrightarrow{~\texttt{asgn}~}}
\newcommand{\phiarrow}{\xrightarrow{phi}}
\newcommand{\controlflowarrow}{\xrightarrow{~\texttt{ctrl}~}}

\makeatletter\if@ACM@journal\makeatother
%% Journal information (used by PACMPL format)
%% Supplied to authors by publisher for camera-ready submission
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2017}
\acmMonth{1}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\else\makeatother
%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
\acmYear{2017}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission
\setcopyright{none}             %% For review submission
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
%\citestyle{acmauthoryear}  %% For author/year citations
%\citestyle{acmnumeric}     %% For numeric citations
%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
                            %% sorting of references within a single citation;
                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
                            %% rendered as [14,5,2] rather than [2,5,14].
%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
                            %% compression of sequential references within a
                            %% single citation;
                            %% e.g., \cite{Baker12,Baker14,Baker16}
                            %% rendered as [2,3,4] rather than [2-4].

\begin{document}

%% Title information
\title[Short Title]{Catching fire: Undefined behaviour for equational reasoning of lazy languages with strictness annotations}
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\subtitle{Subtitle}                     %% \subtitle is optional

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}
}
\email{first2.last2@inst2b.org}         %% \email is recommended

\begin{abstract}
% An abstract should consist of six main sentences:
%  1. Introduction. In one sentence, what’s the topic?
%  2. State the problem you tackle.
%  3. Summarize (in one sentence) why nobody else has adequately answered the research question yet.
%  4. Explain, in one sentence, how you tackled the research question.
%  5. In one sentence, how did you go about doing the research that follows from your big idea.
%  6. As a single sentence, what’s the key impact of your research?

% 1
We propose a new style of semantics
for non-strict languages that provide access to many optimizations that are
currently unsound for lazy languages.
% 2 
Lazy languages, while providing equational reasoning, are unable to reason
effectively about \emph{strictness annotations}. On the other hand, most strict
languages are impure, and thus unable to provide equational reasoning due to
the presence of side effects.
% 3
Research on optimizing non-strict languages has focused on optimizing away
non-strictness with demands, and efficient lowering of strictness onto modern
hardware.
% 4
We take a different approach: we propose to use undefined behaviour judiciously
while defining the semantics of our language to enable many key optimizations that
are possible in the strict world, that are not possible in the non-strict world.
We model \emph{divergent computations as undefined behaviour}, giving the
compiler far more freedom to optimize and reorder a mix of strict and non-strict
code.
% 5 
We first consider the various deficiencies of introducing an explicit "bang pattern"
or "strictness annotation" as is done within the Glasgow Haskell compiler today.
Then we demonstrate how setting divergent computations to be undefined behaviour 
cleanly enables the compiler to optimize these programs.
% 6
This modelling of divergence as undefined behaviour enables us to write an
\emph{optimizing} compiler that can reason about a mix of strictness and non
strictness. Furthermore, it enables us to modularly add non-strictness into
an existing strict language. We demonstrate this with a prototype optimizing
functional language called Lizzy. Lizzy uses MLIR as a backend. We introduce
a new dialect \lz in MLIR for modularly expressing non-strictness.  As a benchmark,
we consider a shootout between GHC's optimization of vector and matrix
computations, and \lz's ability to optimize the same computations.
% (http://www.easterbrook.ca/steve/2010/01/how-to-write-a-scientific-abstract-in-six-easy-steps/)
\end{abstract}

% Only add ACM notes and keywords in camera ready version
% Drop citations and footnotes in draft and blind mode.
\ifx\paperversion\paperversioncamera
%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3}  %% \keywords is optional

\else
\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\fi

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle
\ifx\grammarly\grammarlyon 
\onecolumn 
\else 
\fi

\section{Introduction}

% borrowed from : https://arxiv.org/abs/2003.00532 (Uday's note on MLIR)
Lazy languages such as Haskell
offer many benefits, chief of which is the ability to reason
equationally about programs. For performance, considerations,
the prototypical lazy language, Haskell, provides ``bang patterns'' to
mark strictness. Herein lies the trouble; these ``bang patterns'' do not
permit equational reasoning for the user. Worse, they form a sequence point
in the otherwise pure semantics, taking away most of the optimization freedom
from the compiler. We propose to remedy the situation by making judicious use
of \emph{undefined behaviour semantics}. While undefined behaviour is often
reviled in discourse, it is in fact a major reason why C and C++ can be optimized
as well as they can; defining certain rare or impossible conditions as undefined
behaviour provides the compiler a great deal of freedom by being able to ignore
these cases. We choose to mark \emph{divergence as undefined behaviour}. This
solves the \emph{correctness problem} in reordering strict computations; The
only thing our compiler (Lizzy) will reason about is performance when reordering
strict computations. Furthermore, this is not too insane a proposal; After all,
the C standard also defines side-effect free infinite loops as undefined behaviour.
Next, we show a variety of examples where this simple extension to the language
semantics permits far easier reasoning about mixing strictness and non-strictness.
We then put this to the test, by extending MLIR with a new non-strict IR, called
\lz. Note that the undefined behaviour semantics are critical in order to
reason effectively about the mix of strict and non strict behaviour we wish
to model within MLIR. Finally, to validate our approach, we consider a series
of benchmarks where our programming language, Lizzy outperforms similar programs
in Haskell compiled by the Glasgow Haskell Compiler.

% \begin{figure}
% % Link to figure
% %
% % https://docs.google.com/drawings/d/1juKp43D3rLC-luBQPwQZ_wCnDK2S_6C1k6USV0wKE0g/edit?usp=sharing
% \includegraphics[width=\columnwidth]{images/overview.pdf}
% \caption{Our key idea visualized}
% \end{figure}

% \grosser{Always state your contributions explicitly: (A) this makes it
% easy for the reader to understand what novelty is presented, and (B)
% these contributions help you to focus. In particular, the objectie of
% the remaining paper should be to support the claims stated here.
% }
Our contributions are:                                          
\begin{itemize}
\item A formal semantics for bang patterns in the Haskell language.
\item A study of where bang patterns inhibit GHC's ability to reason about strictness.
\item A novel yet simple extension to a lazy language semantics that remove this
      inability of GHC (and lazy languages in general) to reason about strictness.
\item An implementation of a prototype lazy language (Lizzy) that chooses this
      approach pervasively.
\item A new dialect (\lz) for MLIR to encode these semantics, and provide modular
      non-strictness.
\end{itemize}

\section{Strictness Annotations: The Bad And The Ugly}

We first motivate why strictness annotations inhibit the compiler
from being able to optimize programs. Consider the program in
\autoref{fig:lazy-loop}, where \texttt{out} declares
a binding \raw{x = loop} and proceeds to not use it. This allows GHC
to optimize out the program into \raw{out = 42}. On the
other hand, consider  \autoref{fig:strict-loop} where \raw{out}
declares  \emph{strict} binding \raw{!x  = loop} and proceeds to not
use it.

\begin{lstlisting}[language=haskell,caption=unoptimized \raw{out} and optimized \raw{outOptimized}]
{-# LANGUAGE BangPatterns #-}
loop, out, out' :: Int
loop = loop;
out = let x = loop in 42
outOptimized = 42
main :: IO ()
main = do print out; print out'
\end{lstlisting}

\begin{lstlisting}[language=haskell, caption=unoptimized \raw{outStrict} with an incorrect optimization in \raw{outStrictOptimized}]
{-# LANGUAGE BangPatterns #-}
loop, out, out' :: Int
loop = loop;
outStrict = let !x = loop in 42
outStrictOptimized = 42 -- INCORRECT TRANSFORM
main :: IO ()
main = do print out; print out'
\end{lstlisting}
% \caption{An example of how strictness annotations break equational reasoning}

\begin{lstlisting}[language=haskell, caption=Original haskell program]
{-# LANGUAGE BangPatterns #-}
out b = let large = unknown 10 in 
  if False then large else 42
out' b = let !large = unknown  10 in 
  if False then large else 42
main :: IO ()
main = do print out; print out'
\end{lstlisting}

\begin{lstlisting}[language=Haskell, caption=Correct optimization that is incorrect with a bang pattern]
{-# LANGUAGE BangPatterns #-}
loop, out, out' :: Int
loop = loop;
out b= if b then 
  let large = unknown 10 in large else 42
-- | incorrect. If `large` had been divergent, 
-- then `out` no longer
-- exhibits divergence
out' b = if b then 
  let !large = unknown 10 in large else 42

main :: IO ()
main = do print out; print out'
\end{lstlisting}
% \caption{An example of how strictness annotations disallow floating a \texttt{let}
%          down into the use-site, inhibiting optimization}

\section{Undefined Behaviour For Divergence?}

As we have shown, the existence of bang patterns obviates equational reasoning.
We are now in a bind: we can choose to forego bang patterns, thereby
regaining equational reasoning with a loss of performance. This is not a 
very reasonable suggestion, as bang patterns are critical for controlling
evaluation ordering, and ensuring that we do not build up thunks.

We first propose a simple, outrageous, solution:  Ban all divergence.
We define divergence to be \emph{undefined behaviour}. In essence, this
allows the compiler to assume that \emph{no value} is divergent, for if a
value were divergent, then \emph{anything} can happen. We analyze the
merits and deficiencies of this proposal, and then provide a more
refined version of the above solution.

% \begin{figure}[ht]
% \begin{minted}{hs}
% {-# LANGUAGE BangPatterns #-}
% loop, out, out' :: Int
% outStrict = let !x = loop in 42
% -- ==If x were not divergent==, 
% --   it is safe to set out' to 42.
% -- ==If x were divergent==
% --   It is undefined behaviour. Anything can
% --   happen, including evaluating out' to 42.
% outStrictOptimized = 42 
% main :: IO ()
% main = do print out; print out'
% \end{minted}
% \caption{An example of how defining divergence as UB allows continued
%          equational reasoning with bang patterns.}
% \label{fig:strict-loop-ub}
% \end{figure}

Consider the program in \autoref{fig:strict-loop}. Here, according
to our semantics, we define \emph{divergence as undefined behaviour}.
Interpreted according to these semantics, here's how we legally
allow the transformation \raw{outStrictOptimized = 42}:

\begin{itemize}
    \item Since \raw{x} is forced, we must evaluate \raw{loop} into weak head normal form.
    \item \raw{loop} does not have a weak head normal form. This is a divergent computation.
    \item Since we have divergence, \emph{anything} can happen [we can refine $\bot$ into a
    value that is greater than $\bot$].
    \item We choose to refine $\bot$ into $42$ as $\bot \sqsubset 42$.
\end{itemize}

%\begin{figure}
%\begin{minted}{hs}
\begin{lstlisting}[language=haskell, caption=An example of how defining divergence as UB is too agressive]
{-# LANGUAGE BangPatterns #-}
-- | never evaluates to a value.
main = putStrLn "42"; main
mainOptimized = return () 
 -- | legal if divergence is UB.
\end{lstlisting}
%\end{minted}
%\label{fig:io-loop}
%\caption{An example of how defining divergence as UB is too agressive}
%\end{figure}

On the other hand, this definition is \emph{far too agressive}. Consider
the program in \autoref{fig:io-loop}. Here, main prints $42$ forever,
never returning the final \raw{IO ()} that it is expected to. Thus,
we can \emph{optimize this program} into \raw{mainOptimized = return ()}
if divergence is undefined behaviour. This is not what we want! We wish
to allow programs which \emph{make progess without halting}, while
still being able to exploit undefined behaviour. Thus, we must change our
semantics to adapt to this need.

\section{Progressive operational semantics}

Naively speaking, one maybe tempted to summarize our approach as
"make divergence undefined". While this is appealing, this is both too
strong, and too naive. For example, such a definition would allow us to
optimize away programs that \emph{progress} without \emph{terminating}.
A tiny example of such a program, which prints ``42'' forever is
given in \autoref{fig:print-forever}

% \begin{figure}[ht]
% \begin{minted}{hs}
\begin{lstlisting}[language=haskell, caption=A program that makes progress without terminating.  We wish to preserve such program semantics]
action :: IO (); action = putStrLn "42";
main :: IO (); main = action >> main
\end{lstlisting}
% \end{minted}
% \label{fig:print-forever}
% \caption{A program that makes progress without terminating.  We wish to preserve such program semantics}
% \end{figure}

% \begin{figure}[ht]

%\begin{subfigure}[b]{0.48\textwidth}
%\begin{minted}{hs}
\begin{lstlisting}[language=haskell]
{-# LANGUAGE BangPatterns #-}
loop :: Int; loop = loop
ex1 = let x = loop in 42
ex2 = let !x = loop in 42
ex3 = let !x = loop in x
-- | New program
{-# LANGUAGE BangPatterns #-}
loop :: Int; loop = loop
ex1 = 42
ex2 = 42
ex3 = let !x = loop in x
ex3' = loop -- equivalent semantics
\end{lstlisting}
% \end{minted}
% \subcaption{Maxmially transformed programs with our proposed semantics}
% \end{subfigure}
% \caption{Programs that exhibit a range of behviours, clarifying what values we do and do not eliminate}
% \end{figure}

We call our semantics \emph{progressive}, which tries to make as much progress
as possible before getting stuck at a stuck term.
\section{Formal Semantics Of Bang Patterns in Haskell}

There is currently no formal definition of bang patterns in the Haskell language.
The semantics are defined based on how the Glasgow Haskell Compiler chooses
to manipulate these strictness annotations. So wherever we have previously argued
that it would be ``incorrect'' for GHC to perform these transforms, we have been
arguing from the position of:

\begin{itemize}
\item This is what GHC does today, and that is the current state of definition for bang patterns.
\item Were GHC to make the above transformation, we believe it would violate the
      intuitions that many people have about how GHC \emph{ought} to transform code.
      \sid{I asked bgamari who works on GHC full-time, he said that GHC does not perform the above transform most likely.
           Similarly, other friends (carter, davean) who commit into GHC also said likely not.}
\end{itemize}

To begin our discussion, we first formalize the semantics of bang patterns.
\textbf{TODO: fill in semantics here.}
\section{\lz: Its Syntax, Type system, semantics}
\subsection{A high level overview of \mlir}

\begin{figure*}
\begin{minted}{text}
toplevel := func | global
func := <fn-name> formal-param-list -> <ret-type> <region>
region := list [bb]
bb := <bb-name> formal-param-list : list inst; terminator-inst
inst := retval "=" <op-name> arg-list
terminator-inst := std.return(<name>) | br <name> | condbr <name> <bb1> <param-list> <bb2> <param-list>
\end{minted}
\caption{The \std dialect syntax}

\end{figure*}

\begin{figure*}
\vspace{3em}

\begin{tabular}{l l}
\begin{minipage}[t][1cm][b]{0.5\textwidth}
\inference[assign-cst]
{
}
{
  (\texttt{\%vid =}~ \texttt{std.constant } const, R) ~\assignarrow~R[\texttt{\%vid} \mapsto const]
}
\end{minipage}

&

\begin{minipage}[t][1cm][b]{0.5\textwidth}
\inference[assign-fncall]
{
  R[x_1] = v_1\quad R[x_2] = v_2\quad \dots\quad R[x_n] = v_n \\
  \denotation{f}(v_1, v_2, \dots, v_n) = y \\
}
{
  (\texttt{\%vid =}~ \texttt{f(}x_1, x_2, \dots, x_n \texttt{)}, R) ~\assignarrow~R[\texttt{\%vid} \mapsto y]
}
\end{minipage}
\end{tabular}

% =========================

\vspace{3em}
\begin{tabular}{l l}
\begin{minipage}[t][1cm][b]{0.5\textwidth}
\inference[cbr-left]
{
    % P[pc] = \texttt{cond\_br \%c $\caret$ bb1(...) \^bb2(...)} \quad
    P[pc] = \raw{condbr c ^bb1(xs) ^bb2(ys)} \\
    P[\raw{^bb1}] =  \raw{^bb1(ls)} \\
    R[c] \ne 0
}
{
    (pc, R) ~\controlflowarrow~(pc', R[ls \mapsto R[xs])
}

\end{minipage}
&
\begin{minipage}[t][1cm][b]{0.5\textwidth}

\inference[cbr-right]
{
    P[pc] = \texttt{br}~ c~ l(v_1)~ r(v_2) \quad
    P[r] = pc' \quad
    R[c] = 0
}
{
    (pc, R) ~\controlflowarrow~(pc', R[r.input \mapsto R[v_2])
}
\end{minipage}
\end{tabular}

% =========================

\inference[br]
{
    P[pc] = \texttt{br}~ n(v) \quad
    P[n] = pc'
}
{
    (pc, R) ~\controlflowarrow~(pc', R[n.input \mapsto R[v])
}

\caption{The operational semantics of the \std dialect}
\end{figure*}

\subsection{A high level overview of \lz}
We extend the basic \std dialect modestly, by adding primitives that can 
describe non-strictness. We first introduce a primitive known as \lzthunk,
which is responsible for creating non-strict thunks. 
To represent lazy function application, we introduce a new operation called \lzap,
which is the lazy sibling of \stdcall. This creates a thunk, which when evaluated,
invokes the function with the given arguments. To force
a thunk, we introduce \lzforce, which receives a thunk as input and returns
the forced value of the thunk as output. In effect, we have an encoding of
administrative normal form as (ANF) an MLIR dialect.

To represent algebraic data types, which enable reasoning of control flow 
and are the primary mode of abstraction in most functional languages, as
well as newer imperative languages such as Rust, we introduce the \construct
and \case operators, which are the introduction and elimination forms of
structured data in the dialect.

\begin{figure*}[hbt!]
\begin{minted}{text}
thunk(x): T -> Thunk<T>
ap(f, v1, v2, ..., vn): ((T1,...Tn) -> R) x T1 x ... Tn -> Thunk<R>
force(x): Thunk<T> -> T
construct(ConsName, v1, ..., vn): Symbol x T1 x T2 ... x Tn -> ADT<T>
case(x: T) of ... : ADT<T> -> R
\end{minted}
\caption{Syntax of \lz extensions}
\end{figure*}

\begin{figure*}
\vspace{3em}
\begin{tabular}{l l}
\begin{minipage}[t][1cm][b]{0.5\textwidth}
\inference[asgn-constructor]
{
   P[pc] = \\
    \quad \texttt{\%vid = lz.construct(@ConsName, \%x1, ..., \%xn)} \\
   P[\texttt{\%x1}] = v_1, \quad \dots, \quad P[\texttt{\%xn}] = v_n \\
}
{
  R[\texttt{\%vid} \mapsto \texttt{constructor(@ConsName, $v_1, \dots, v_n$)}]
}
\end{minipage}
&
\begin{minipage}[t][1cm][b]{0.5\textwidth}
\inference[asgn-thunk]
{
   P[pc] = \texttt{\%vid = lz.ap(\%fref, \%x1, ..., \%xn)} \\
   P[\texttt{\%f1}] = \texttt{ref(@fn)} \\
   P[\texttt{\%x1}] = v_1, \quad \dots, \quad P[\texttt{\%xn}] = v_n \\
}
{
  R[\texttt{\%vid} \mapsto \texttt{thunk(@fn, $v_1, \dots, v_n$)}]
}
\end{minipage}
\end{tabular}

% ================

\vspace{5em}
\begin{tabular}{l l}
\begin{minipage}[t][1cm][b]{0.5\textwidth}
\inference[asgn-case]
{
   P[pc] = \texttt{\%vid = lz.case(\%x, @Cons\textsubscript{1}, r\textsubscript{1},}\dots 
                    \texttt{, @Cons\textsubscript{N}, r \textsubscript{N})} \\
   \texttt{@Cons}_i = \texttt{@XCons} \\ 
   P[\texttt{\%x}] = \texttt{construct(@XCons, $v_1, \dots, v_n$)} \\
   r_i = \raw{ { ^entry(arg1, ..., argn): ... } }
}
{
  % decide on how to represent region execution
  (pc, R) \controlflowarrow (P[r_i], R[\texttt{arg}_i \mapsto v_i])
}
\end{minipage}
&
\begin{minipage}[t][1cm][b]{0.5\textwidth}
\inference[asgn-force]
{
   p[pc] = \texttt{\%vid = lz.force(\%thnk)} \\
   p[\texttt{\%thnk}] = \texttt{thunk(@fn, $v_1, \dots, v_n$)} \\
   \denotation{@fn}(v_1, v_2, \dots, v_n) = y
}
{
  r[\texttt{\%vid} \mapsto y]
}
\end{minipage}
\end{tabular}
\caption{Operational semantics of \lz  extensions}
\end{figure*}
\textbf{TODO: this is broken, because it doesn't scope properly. We need to describe \lzthunk
as building a thunk \emph{in memory/on the heap}, and the value stored in the register file
is the pointer to this thunk}.

\subsection{Purifying \lzforce}

Let us for a moment assume that divergence is not undefined behaviour. Now consider
the code in \autoref{fig:unused-divergent-variable-lz}. The variable called
\texttt{\%loopv} is unused. However, because divergence \texttt{\%loopv} is
divergent, it is incorrect to eliminate the variable definition of 
\texttt{\%loopv}, for the divergence of \texttt{loop} occurs when \texttt{\%loopt} is forced;
That is, when the instruction \texttt{\%loopv = lz.force(\%loopv)} is executed.
This implies that the instruction \lzforce is side-effecting.

If we wished to regain purity of \lzforce and leverage the power of
SSA, then we must be allowed to eliminate a call to \lzforce if the
value is unused. This is exactly the same requirement as being able to 
equationally reason with bangs in a \texttt{let} binding. In that use-case,
we wished for undefined behaviour semantics for ease-of-reasoning. Here, we wish
for undefined behaviour semantics for ease-of-optimization.

\begin{lstlisting}[language=c++, caption=The SSA encoding of a lazy program with divergence in the \lz dialect.  Note the unused divergent variable \texttt{\%loopv}]
// loop = loop
lz.func @loop () -> !lz.adt<@Box> {
  %outt = lz.ap(@loop) : !lz.thunk<lz.adt<@Box>> 
  %outv = lz.force(@loop) : !lz.adt<@Box>
  lz.return(%outv) : !lz.value
}

// main = let !x = loop in 42
lz.func @main() -> !lz.adt<@Int> {
  %loopt = lz.ap(@loop) 
    : !lz.thunk<lz.adt<@Box>>
  %loopv = lz.force(%loopv) // <- unused %loopv
    : !lz.adt<@Box>
  %i = std.constant 42 : i64
  %retv = lz.construct(@Int, i)
  lz.return(%retv)
}
\end{lstlisting}
% \caption{The SSA encoding of a lazy program with divergence in the \lz dialect.
%          Note the unused divergent variable \texttt{\%loopv}}
% \label{fig:unused-divergent-variable-lz}
% \end{figure}

%\begin{figure}[hbt]
%\begin{minted}{text}
\begin{lstlisting}[language=c++, caption=Under the assumption that divergence is UB we can eliminate the call to \lzforce. This regains SSA semantics.]
// loop = loop
lz.func @loop () -> !lz.adt<@Box> {
  %outt = lz.ap(@loop) : !lz.thunk<lz.adt<@Box>> 
  %outv = lz.force(@loop) : !lz.adt<@Box>
  lz.return(%outv) : !lz.value
}

// main = let !x = loop in 42
lz.func @main() -> !lz.adt<@Int> {
  // %loopt = lz.ap(@loop) 
  //  : !lz.thunk<lz.adt<@Box>>
  // %loopv = lz.force(%loopv) 
  //   : !lz.adt<@Box> 
  %i = std.constant 42 : i64
  %retv = lz.construct(@Int, i)
  lz.return(%retv)
}
\end{lstlisting}
% \caption{Under the assumption that divergence is UB, we can eliminate the call to
%          \lzforce. Thus, this assumption allows us to regain SSA semantics.}
% \label{fig:unused-divergent-variable-lz-eliminated}
% \end{figure}

\subsection{Purifying \lzconstruct}

Consider the program in \autoref{fig:unused-construct}. For us to eliminate
the call to \texttt{lz.construct}, we need to be sure that memory allocation
has no side effects. In our case, since we assume that allocations are not
visible to the user in terms of their implementation details (eg. The address
of the pointer of the block of memory), we can safely remove the value
\texttt{\%unused = lz.construct(@Box)}  as \texttt{\%unused} is unused.

On the flipside, consider a program where we want to \emph{copy} the 
instruction \texttt{lz.construct}. If we were not guaranteed the existence
of a garbage collector, then this would not be pure; we would have created a
chunk of memory that is never freed. \textbf{TODO: Is there a good example where we 
want to copy a value for optimization? Something like, copy an SSA value both
into a loop and outside or something?}

\begin{lstlisting}[caption=An unused called to \lzconstruct]
lz.func() -> !hask.adt<@Box> {
  %unused = lz.construct(@Box)
  %one = std.constant 1 : i64
  lz.ret(%one) : i64
}
\end{lstlisting}

\section{Aiding the performance programmer with explicit forcing}

\begin{lstlisting}[language=haskell, caption=Haskell program with two datatypes  \texttt{Lazy} and \texttt{Strict}
with the same shape of data but different strictness annotations.]
{-# LANGUAGE BangPatterns #-}
{-# LANGUAGE UnboxedTuples #-}
module Foo(boxLazyAndStrict) where
-- | A large datatype
data Box = Box Int Int Int Int Int
data Lazy = MkLazy Box
data Strict = MkStrict !Box

data BoxPair = MkBoxPair !Box !Box

boxLazyAndStrict :: Lazy 
  -> Strict 
  -> BoxPair
boxLazyAndStrict lz str =
  case lz of
   MkLazy lbox  ->
       case str of
         MkStrict strbox ->
           MkBoxPair lbox strbox
\end{lstlisting}

\begin{lstlisting}[language=haskell,caption=\ghc \core corresponding to the  above Haskell program  where there is no indication
         that \texttt{strbox} is a value while \texttt{lbox} is a thunk]
boxLazyAndStrict :: Lazy -> Strict -> BoxPair
 ...
 Tmpl= \ (lz_aub :: Lazy)
         (str_auc :: Strict) ->
         case lz_aub of { MkLazy lbox_aud ->
         case str_auc of { MkStrict strbox_aue ->
         case lbox_aud of dt_Xve [Occ=Once]
         { Box _ _ _ _ _  ->
         Foo.MkBoxPair dt_Xve strbox_aue
         }}}}]
\end{lstlisting}

In Haskell, Core and STG, all of GHC's intermediate representations, there is
no explicit description of the act of forcing a thunk. The \texttt{case} construct
is responsible for \emph{both} forcing and case analysis. This is a fine state
of affairs when we are in a language where values are forced right before they
are used --- that is, the language is \emph{truly lazy}. However, Haskell currently
supports bang patterns in defining data constructors, which violates this
assumption. We begin with an example of a simple program with two datatypes,
called \texttt{Lazy} and \texttt{Strict}, where the both contain the lazy data
of type \texttt{Box}. \texttt{Strict}, to its name, is actually strict
in its \texttt{Box} field. However, this is completely opaque on reading the
Core: We need to infer from the facts:
\begin{itemize}
\item because \raw{lbox_aud} is \texttt{case}'d on
\item The result of the aforementioned case is called \raw{dt_Xve}.
\item The case is followed by the pointless match \raw{(Box _ _ _ _ _)}.
\item Finally, the \raw{(Foo.MkBoxPair dt_Xve ...)} is built.
\item Thus, the only reason \raw{dt_Xve} exists is to represent the forced \raw{lbox_aud}.
\item So, the value \raw{lbox_aud} is forced into the value \raw{dt_Xve}
\end{itemize}

These infelicities, while slight, add up rapidly when attempting to optimize
more complicated Haskell programs. On the other hand, consider our encoding
in \lz. Since we separate the concept of \emph{forcing} a value from \emph{pattern matching},
we can represent the force necessary as a \texttt{lz.force}, without needlessly
introducing a pattern match. This improves the readability of the source, and
more directly expresses the semantics we are interested in.  While GHC Core
is an amazing intermediate representation for analysis and transformation, it
is sub-par for \emph{consumption by the end-user}. Though the source language has
gotten more complex, \core has changed very little; This is a source of
strength (as the optimizer needs to only reasons about a very minimal language).
It is also a source of obfuscation, as many overlapping ideas are represented
by the same construct: in this case, that of forcing, and that of pattern
matching.

\begin{lstlisting}[language=llvm, caption=\lz dialect encoding of the above program where there is a clear indication of the thunk being forced]
lz.func(%lz: !lz.adt<@Lazy>, 
        %str: !lz.adt<Strict>) {
  %lboxt = lz.extractvalue(%lz, 0)
     : !lz.thunk<!lz.adt<@Box>>
  %strboxv = lz.extractvalue(%str, 1)
     : !lz.adt<@Box>
  %lboxv = lz.force(%lboxt) # FORCE
    : !lz.adt<@Box>
  %retval = lz.construct(@MkBoxPair, 
                         %lboxv, %strboxv)
    : !lz.adt<@BoxPair>
  lz.return(%retval)
}
\end{lstlisting}

\section{Worker wrapper by local rewrites: A first stab}
In this section, we consider a small example program, which we will explain
how to perform a classical transform (the worker-wrapper transform) purely
by using local rewrites. Traditionally, such a transformation is performed
by first performing a \emph{demand analysis} whose results are used
to drive the rewrite.

We explore how this transformation can be achieved \emph{without} the need
for demand analysis by performing \emph{local rewrites}. This provides
both a simpler manifestation of the algorithm, as well as a potentially 
faster implementation, as the MLIR infrastructure is capable of performing
rewrites in parallel.
% \begin{figure}[htb!]
%\begin{minted}[linenos]{cpp}
{\footnotesize
\begin{lstlisting}[caption=initial source code]
SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    int ihash = casedefault(icons);
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        SimpleInt siprev = SimpleInt(prev);
        Thunk<SimpleInt> siprev_t = thunkify(siprev);
        SimpleInt f_prev_v = apStrict(f, siprev_t);
        return f_prev_v;
    }
}

int main() {
    printf("%d\n", f(thunkify(SimpleInt(1))).v);
}
\end{lstlisting}
}
%\end{minted}
%\caption{Step 0: The source code we optimize using worker wrapper}
%\end{figure}

%\begin{figure}[htb!]
%\begin{minted}{cpp}
{\footnotesize
\begin{lstlisting}[caption=Step 1: outlining everything after the initial \lzforce into \texttt{f2}]
SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        SimpleInt siprev = SimpleInt(prev);
        Thunk<SimpleInt> siprev_t = thunkify(siprev);
        SimpleInt f_prev_v = apStrict(f, siprev_t);
        return f_prev_v;
    }
}

SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    SimpleInt ret = apStrict(f2, icons);
    return ret;
}
\end{lstlisting}
}
%\caption{Step 1: outlining everything after the initial \lzforce into \texttt{f2}}
%\end{figure}

%\begin{figure}[htb!]
%\begin{minted}{cpp}
\begin{lstlisting}[language=c++, caption=Step 2: Removing the strict application]
SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        SimpleInt siprev = SimpleInt(prev);
        // Thunk<SimpleInt> siprev_t = thunkify(siprev);
        // SimpleInt f_prev_v = apStrict(f, siprev_t);
        Thunk<SimpleInt> f_prev_v = apStrict(f2, siprev);
        return f_prev_v;
    }
}

SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    SimpleInt ret = apStrict(f2, icons);
    return ret;
}
\end{lstlisting}
%\end{minted}
%\caption{Step 2: Removing the strict application}
%\end{figure}

%\begin{figure}[htb!]
{\footnotesize
\begin{lstlisting}[language=c++, caption=Step 3: outline everything after casedefault into f3]
SimpleInt f3(int icons) {
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        SimpleInt siprev = SimpleInt(prev);
        Thunk<SimpleInt> f_prev_v = apStrict(f2, siprev);
        return f_prev_v;
    }
}

SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    SimpleInt ret = apStrict(f3, ihash);
    return ret;
}

SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    SimpleInt ret = apStrict(f2, icons);
    return ret;
}
\end{lstlisting}
}
%\caption{ Step 3: outline everything after casedefault(icons) into f3}
%\end{figure}

%\begin{figure}[htb!]
{\footnotesize
\begin{lstlisting}[language=c++, caption=replace call \texttt{f2(SimpleInt(prev))} to \texttt{f3(prev)}]
SimpleInt f3(int icons) {
    if (ihash <= 0) {
        return SimpleInt(42);
    } else {
        int prev = ihash - 1;
        Thunk<SimpleInt> f_prev_v = apStrict(f3, prev);
        return f_prev_v;
    }
}
SimpleInt f2(SimpleInt icons) {
    int ihash = casedefault(icons);
    SimpleInt ret = apStrict(f3, ihash);
    return ret;
}
SimpleInt f(Thunk<SimpleInt> i) {
    SimpleInt icons = force(i);
    SimpleInt ret = apStrict(f2, icons);
    return ret;
}
\end{lstlisting}
}
%\caption{Step 4: replace call f2(SimpleInt(prev)) to f3(prev)}
%\end{figure}

%\begin{figure}[htb!]
%% ! {\footnotesize
%% ! \begin{lstlisting}[language=c++]
%% ! SimpleInt f3(int icons) {
%% !     if (ihash <= 0) {
%% !         return SimpleInt(42);
%% !     } else {
%% !         int prev = ihash - 1;
%% !         Thunk<SimpleInt> f_prev_v = apStrict(f3, prev);
%% !         return f_prev_v;
%% !     }
%% ! }
%% ! SimpleInt f(Thunk<SimpleInt> i) {
%% !     SimpleInt icons = force(i);
%% !     int ihash = casedefault(icons);
%% !     SimpleInt ret = apStrict(f3, ihash);
%% !     return ret;
%% ! }
%% ! \end{lstlisting}
%% ! }
%\caption{Step 5: Inline f2 into f. This gives us the worker(f3) and wrapper(f)}
%\end{figure}

\section{Worker wrapper by local rewrites: Non-tail-calls}

Currently implementing the constructor simplification

\section{Worker wrapper by local rewrites: Sum types}

\section{Worker wrapper by local rewrites: Sum types + non-tail-calls}

\section{Demand analysis by rewrites: The full algorithm}

\begin{figure*}
\vspace{3em}
{\footnotesize
\begin{tabular}{l l}
\begin{minipage}[t][1cm][b]{0.5\textwidth}
% \begin{subfigure}[b]{0.48\textwidth}

\inference[ForceOfKnownAp]{
\texttt{\%x = ap(\%f, \%v1, ..., \%vn): !lz.thunk<T>} \\
\texttt{\%y = force(\%x): T}
}{
\texttt{\%y = \%f(\%v1, ..., \%vn): T}
}
\subcaption{force of a known function application: remove laziness}
\end{minipage}

&

\begin{minipage}[t][1cm][b]{0.5\textwidth}
\inference[ForceOfThunkify]{
\texttt{\%x = thunkify(\%v) : !lz.thunk<T>} \\
\texttt{\%y = force(\%x): T}
}{
\texttt{\%y = \%x}
}
\subcaption{force of a thunk: remove laziness}
\end{minipage}
\end{tabular}
}

% =========================
\vspace{20em}

\begin{tabular}{l l}
\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize

\inference{
% foo
\texttt{func f(\%x1, ... \%txi: thunk<T>, ..., \%xn) \{} \\
\texttt{  \%xi = force(\%txi)} \\
\texttt{  \%zi = ... ; \%tzi = thunkify(\%zi)} \\
\texttt{  f(\%y1, ..., \%tzi, ..., \%yn)} \\
\texttt{\}}
}{
\texttt{func f\_strict\_i(\%x1, ... \%xi: T, ..., \%xn) \{} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_strict\_i(\%y1, ..., \%zi, ..., \%yn)} \\
\texttt{\}} \\
\texttt{func f(\%x1, ... \%txi: thunk<T>, ..., \%xn) \{} \\
\texttt{   \%xi = force(\%txi)} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_strict\_i(\%y1, \%y2, ..., \%zi, ..., \%yn)} \\
\texttt{\}}
}
} % end footnotesize
\subcaption{outlining recursive call that is immediately forced}
\end{minipage}
&
\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize

\inference{
\texttt{func f(\%x1, ... \%wrapxi: @ADT, ..., \%xn) \{} \\
\texttt{   \%xi = case (\%wrapxi) [\@Wrapper -> \string^ entry(\%v) \{ return \%v \}]} \\
% \texttt{   \%zi = ...}
% \texttt{   \%wrapzi = construct(@Wrapper, \%zi)}
% \texttt{   f(\%y1, ..., \%wrapzi, ..., \%yn)}
\texttt{\}}
}{
\texttt{func f\_work\_i(\%x1, ... \%xi: T, ..., \%xn) \{} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_strict\_i(\%y1, \%y2, ..., \%zi, ..., \%yn)} \\
\texttt{\}} \\
\texttt{func f(\%x1, ... \%txi: thunk<T>, ..., \%xn) \{} \\
\texttt{   \%xi = case (\%wrapxi) [\@Wrapper -> \string^ entry(\%v) \{ return \%v \}]} \\
\texttt{   \%zi = ...} \\
\texttt{   f\_work\_i(\%y1, ..., \%zi, ..., \%yn)} \\
\texttt{\}}
} % end inference
} % end footnotesize
\subcaption{outlining of recursion of a monovariant wrapper}
\end{minipage}
\end{tabular}

\vspace{10em}

\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize
\inference{
\texttt{\%x = constructor(@Constructor, \%v1, ..., \%vm)} \\
\texttt{\%y = case \%x} \\
\texttt{      [\@C1 -> \{\string^entry1(\%z11, ..., \%z1n1): ...  \}]} \\
\texttt{      [...]} \\
\texttt{      [\@Ci -> \{\string^entry1(\%zi1, ..., \%zini): ...  \}]} \\
}{
\texttt{Inline \string^entryi with \%zik = \%vk}
}
} % end footnotesize
\subcaption{Case of known constructor: remove indirection}
\end{minipage}

\vspace{10em}

\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize
\inference{
FOO
}{
BAR
}
} % end footnotesize
\subcaption{Inline a function call}
\end{minipage}
\caption{Local rewrites performed to eliminate laziness (1)}
\end{figure*}

\begin{figure*}

\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize
\inference{
FOO
}{
BAR
}
} % end footnotesize
\subcaption{Outline recursive call of constructor that is immediately unwrapped}
\end{minipage}
% OutlineRecursiveApEagerOfConstructorPattern

\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize
\inference{
FOO
}{
BAR
}
} % end footnotesize
\subcaption{Outline pattern matching branches on a function input}
% OutlineCaseOfFnInput
\end{minipage}

\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize
\inference{
FOO
}{
BAR
}
} % end footnotesize
\subcaption{Outline return of constructor}
% outline return of constructor
\end{minipage}

\begin{minipage}[t][1cm][b]{0.5\textwidth}
{\footnotesize
\inference{
FOO
}{
BAR
}
} % end footnotesize
\subcaption{Peel constructor from case}
% peel constructor from case
\end{minipage}

%% % \begin{minted}{text}
%% % Inlining
%% % \end{minted}
\caption{Local rewrites performed to eliminate laziness (2)}
\end{figure*}

\section{Evaluation}

As a baseline, we compare our performance on \nofib test-suite, which is the
test suite that GHC is tested with and performance improvements to GHC 
are reported on. We only consider a representative subset of programs from \nofib;
The full test suite also extensively tests Haskell's semantics of parallelism, 
software transactional memory, and other features that are orthogonal to the
problem of representing and optimizing non strictness.

\section{Related Work}

\subsection{Demand analysis}
\cite{wadler1987projections} ``Projections for demand analysis'' performs
demand analysis computation by modelling demands as projections on the semantic
domain. We use their broad framework, adapted to our setting.

\cite{grafcall}, \cite{breitner2014call} extend the demand analysis with
finer grained information, such as Call Arity analysis.

\subsection{GHC's intermediate representation}

The Glasgow Haskell compiler \cite{jones1993glasgow} is a mature optimizing
compiler for Haskell, whose plugin infrastructure we hook into to  develop a test
set for \lz, and whose demand analysis we use as a benchmark to measure against.

\subsection{Intel Haskell research compiler IR}
The Intel labs Haskell Research compiler \cite{liu2013intel} models Core in a
strict ANF language. The compiler performs demand analysis and abstract
simplification on ANF. The demand analysis is performed using traditional
abstract interpretation techniques. Later, this demand information is used
to interpret the program and perform abstract simplification. (TODO: figure out
details of this step).

It then compiles to an intermediate representation called MIL,
which is a loosely typed CFG based, SSA-lite intermediate representation.  They
represent laziness as heap values, and manipulate the heap. Thus, their
representation and usage of lazy values reasons with memory, instead of
providing value semantics.
MIL also does not have a notion of nested regions. Therefore, MIL extends
the traditional control flow controls with more finer-grained information,
called as \texttt{cut} and \texttt{interproc}.

\subsection{Mixing dataflow analysis and transformation}

This is based on the theory of compositional dataflow analysis
and transformations, \cite{lerner2002composing} 
which proves that we can interleave dataflow analyses and transformations
safely.

Hoopl \cite{ramsey2010hoopl} is the dataflow analysis and transformation library
within GHC. However, GHC does not use this for performing dataflow analysis
over Core, as the Hoopl library is designed to work with a CFG based
intermediate representation, while GHC Core is a typed functional, expression
oriented intermediate representation. While Hoopl is used for certain simplifications
within C-- in GHC, it is not used extensively due to  poor performance characteristics
exhibited by the implementation. (TODO: bench Hoopl on contrived examples)

\subsection{Alternative encodings of laziness}

GRIN  \cite{boquist1996grin} is an alternative intermediate representation
for lazy and strict functional programming languages which explicitly represents
heap manipulation and case analysis. However, it is not SSA based.

\subsection{Theoretical justification for our encoding}

Our encoding is based on call-by-push-value \cite{levy2012call}, which
breaks down call-by-value and call-by-name paradigms into simple primitives. In
our dialect \lz, we expose these primitives to enable a clean mixture of
call by name and call by value. This enjoys the many properties that
are proven to hold for call-by-push-value (TODO: which ones?!)

\subsection{Technology substrate}

MLIR \cite{lattner2020mlir} is a compiler framework for building reusable and
extensible compiler infrastructure.  MLIR aims to address software
fragmentation, improve compilation for heterogeneous hardware, significantly
reduce the cost of building domain specific compilers, and aid in connecting
existing compilers together.

\section{Conclusion}

We provide a baseline implementation of non-strictness for the MLIR framework
which with very minimal extensions, allows us to fully express non strict semantics
while being compatible with the rest of MLIR's strict dialects.
As an experience report, we explore MLIR's strengths and weaknesses at representing
non-strictness within an SSA based framework. We also provide a baseline implementation
of demand analysis, as inspired by Wadler and Hugh's ``projections for demand
analysis''. We evaluate our example programs against the nofib benchmark test suite.
We also provide provocative examples where the potent combination of MLIR's 
loop optimization infrastructure along with our modest extensions allows us to
beat native haskell's performance.

\section{TODO}

\subsection{Performance benchmarking}
\subsection{Correct encoding for mutual recursion}
\subsection{Correct encoding for partially applied functions/closures}

We currently have a \texttt{lz.lambda} but it's basically untested, because
we didn't have many things that used closures. We need to make sure this
works properly, and encodes data correctly.

\subsubsection{Lambdas in GRIN}
GRIN is also a low level IR, so it's useful to recall how this is encoded
within GRIN. To quite the GRIN thesis:

\say{
end. Using hbcc we get well optimised
code in a low level functional style, comparable to for example the Core lan-
guage [PJ96] used by the Glasgow Haskell compiler.  The code is lambda lifted,
i.e., has only super combinators, and most high level Haskell constructions,
like overloading, have been transformed away.
}

\subsubsection{Implementing lambda lifting}
Recall that optimal lambda lifting is $O(n^2)$. I don't know the
algorithm; I'd have to study it before I implement it.

\subsection{Stretch: rewrite based unification}
\subsection{Stretch: rewrite based tabled typeclass generation}
\subsection{Stretch: As static as possible garbage collection}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography

%% why does this not show?
\bibliography{references}

%% Appendix
\appendix
\section{Appendix}

Text of appendix

\end{document}